Towards Automated Refactoring of Code Clones in Object-Oriented Programming Languages\\- Work in Progress -

### 2019-07-01 sattose/paper.tex
Something with relations

Looking at the example in figure \ref{fig:type2}, we see an example of a type 2 clone that poses no harm to the design of the system. Both methods are, except for their matching structure, completely different in functionality. They operate on different types, call different methods, return different things, etc. Having such a method flagged as a clone does not provide much useful information.

@@

To find such a threshold, we measured the amount of clone classes found for a certain number of tokens. We then looked at the amount of nodes (statements/declarations) that these clones span. The resulting chart is displayed in figure \ref{fig:tokenthreshold}.

@@

\subsection{Clone types}
In this chapter we display the differences between clone type 1-3 \cite{roy2007survey} and type 1R-3R as proposed in section \ref{sec:rtypes}. When running our clone detection script over the corpus, we get the results displayed in figure

\begin{figure}[H]
  \caption{Number of cloned declarations/statements.}
    \includegraphics[width=1\columnwidth]{img/TypeResults}
  \label{fig:tokenthreshold}
\end{figure}

@@

Figure \ref{fig:clonecontext} shows an abstract representation of clone classes and clone instances. The relation of clones through inheritance is measured on clone class level: it involves all child clone instances. The location and contents of clones is measured on clone instance level. A clone's location involves the file it resides in and the range it spans (for example: line 6 col 2 - line 7 col 50). A clone instance contents consists of a list of all statements and declarations it spans.

We analyzed the context of clones in a large corpus of open source projects. For these experiments, we used our CloneRefactor tool. These experiments follow the structure of the context: The relation between clone instances is explained, measured and discussed in Sec.~\ref{chap:relationsinstances}; the location of clone instances is explained, measured and discussed in Sec.~\ref{chap:clonelocation}; the content of clone instances are explained, measured and discussed in Sec.~\ref{chap:clonecontents}.

\subsection{Clone detection results}
Currently, we have implemented two clone detection algorithms into CloneRefactor. The first one finds clones by comparing tokens (excluding whitespace, comments and newlines), equal to the definition of type 1 clones in literature~\cite{roy2007survey}. The second algorithm implements our T1R, as explained in Sec.~\ref{sec:type1r}. The differences between the clones found for these algorithms is displayed in Table \ref{table:clonedet}.

\begin{table}[H]
  \begin{center}
  \caption{CloneRefactor clone detection results for the two different algorithms.} \label{table:clonedet}
  \medskip
\begin{tabular}{|l|l|l|}
\hline
 & \textbf{Type 1} & \textbf{T1R} \\ \hline
\begin{tabular}[c]{@{}l@{}}Amount of lines cloned\end{tabular} & 200,362 & 129,519 \\ \hline
\begin{tabular}[c]{@{}l@{}}Amount of statements/\\declarations cloned\end{tabular} & 182,466 & 118,980 \\ \hline
\begin{tabular}[c]{@{}l@{}}Amount of tokens cloned\end{tabular} & 1,582,845 & 973,596 \\ \hline
\end{tabular}
\end{center}
\end{table}

Looking at Table \ref{table:clonedet}, it becomes apparent that the T1R algorithm finds significantly less clones than the type 1 algorithm. This indicates that about a third of the clones have textual equality, but are not actually equal when considering the types of expressions. This makes these clones less suitable for automated refactoring.

@@

\subsubsection{Categorizing Clone Instance Relations}\label{chap:catcloneinstancerelations}
Fontana et al.~\cite{fontana2015duplicated} describe measurements on 50 open source projects on the relation of clone instances to each other. To do this, they first define several categories for the relation between clone instances in object-oriented languages. A few of these categories are shown in Figure \ref{fig:clonerelation}. These categories are as follows:
\begin{enumerate}
  \item \textbf{Same method}: All instances of the clone class are in the same method.
  \item \textbf{Same class}: All instances of the clone class are in the same class.
  \item \textbf{Superclass}: All instances of the clone class are children and parents of each other.
  \item \textbf{Ancestor class}: All instances of the clone class are superclasses except for the direct superclass.
  \item \textbf{Sibling class}: All instances of the clone class have the same parent class.
  \item \textbf{First cousin class}: All instances of the clone class have the same grandparent class.
\item \textbf{Common hierarchy class}: All instances of the clone class belong to the same hierarchy, but do not belong to any of the other categories.
\item \textbf{Same external superclass}: All instances of the clone class have the same superclass, but this superclass is not included in the project but part of a library.
\item \textbf{Unrelated class}: There is at least one instance in the clone class that is not in the same hierarchy.
\end{enumerate}

\begin{figure}[H]
  \caption{Abstract figure displaying some relations of clone classes. Arrows represent superclass relations.}
    \medskip
    \includegraphics[width=1\columnwidth]{img/Relation}
  \label{fig:clonerelation}
\end{figure}
Please note that none of these categories allow external classes (except for ``same external superclass''). So if two clone instances are related through external classes but do not share a common external superclass, it will be flagged as ``unrelated''. The main reason for this is that it is (often) not possible to refactor to external classes.

>>>>

Looking at the example in Fig.~\ref{fig:type2}, we see an example of a type 2 clone that poses no harm to the design of the system. Both methods are, except for their matching structure, completely different in functionality. They operate on different types, call different methods, return different things, etc. Having such a method flagged as a clone does not provide much useful information.

@@

To find such a threshold, we measured the amount of clone classes found for a certain number of tokens. We then looked at the amount of nodes (statements/declarations) that these clones span. The resulting chart is displayed in Fig.~\ref{fig:tokenthreshold}.

@@

\subsection{Clone types}
In this chapter we display the differences between clone type 1-3 \cite{roy2007survey} and type 1R-3R as proposed in section \ref{sec:rtypes}. When running our clone detection script over the corpus, we get the results displayed in Fig.~\ref{fig:typeres}.

\begin{figure}[H]
  \caption{Number of cloned declarations/statements.}
    \includegraphics[width=1\columnwidth]{img/TypeResults}
  \label{fig:typeres}
\end{figure}

In this figure the amount cloned statements per clone type are displayed. The difference between T1R and T1 is small (10.9\%), because most often textually equal code is also functionally equal. The difference between T2R and T2 is bigger (34.7\%), mainly because we allow no variability in used types. T3R and T3 are similar to T2R and T2, because our dataset does not have so many gapped clones for the thresholds used.

We also measured the duration of finding clones by the different clone types. These results are displayed in figure TODO.

@@

Fig.~\ref{fig:clonecontext} shows an abstract representation of clone classes and clone instances. The relation of clones through inheritance is measured on clone class level: it involves all child clone instances. The location and contents of clones is measured on clone instance level. A clone's location involves the file it resides in and the range it spans (for example: line 6 col 2 - line 7 col 50). A clone instance contents consists of a list of all statements and declarations it spans.

@@

\subsubsection{Categorizing Clone Instance Relations}\label{chap:catcloneinstancerelations}
Fontana et al.~\cite{fontana2015duplicated} describe measurements on 50 open source projects on the relation of clone instances to each other. To do this, they first define several categories for the relation between clone instances in object-oriented languages. These categories are shown in Fig.~\ref{fig:clonerelation}. These categories are as follows:
\begin{enumerate}
  \item \textbf{Same method}: All instances of the clone class are in the same method.
  \item \textbf{Same class}: All instances of the clone class are in the same class.
  \item \textbf{Superclass}: All instances of the clone class are children and parents of each other.
  \item \textbf{Ancestor class}: All instances of the clone class are superclasses except for the direct superclass.
  \item \textbf{Sibling class}: All instances of the clone class have the same parent class.
  \item \textbf{First cousin class}: All instances of the clone class have the same grandparent class.
\item \textbf{Common hierarchy class}: All instances of the clone class belong to the same hierarchy, but do not belong to any of the other categories.
\item \textbf{Same external superclass}: All instances of the clone class have the same superclass, but this superclass is not included in the project but part of a library.
\item \textbf{Unrelated class}: There is at least one instance in the clone class that is not in the same hierarchy.
\end{enumerate}

\begin{figure}[H]
  \caption{Abstract figure displaying some relations of clone classes. Arrows represent superclass relations.}
    \medskip
    \includegraphics[width=1\columnwidth]{img/Relation}
  \label{fig:clonerelation}
\end{figure}
